---
editor_options:
  markdown:
    wrap: 72
output: pdf_document
---

# Metadata

    Course:   DS 5100
    Module:   11 R Programming 2
    Topic:    HW on Tidyverse
    Author:   R.C. Alvarado (adapted)
    Date:     6 July 2022

# Student Metadata

    Name:  Rachel Grace
    UVA ID: rg5xm

# Instructions

Perform the tasks below to write the necessary code and include all
solutions.

Read about the Abalone dataset
[here](https://archive.ics.uci.edu/ml/datasets/Abalone).

Grab the `abalone.data` dataset from this URL:

> <https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data>

Hine 1: You can pass the URL directly to `read.csv()`.

Hint 2: there is no header row.
```{r}
library(tidyverse)
```

```{r}
dataset <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data")
```


Note: The instruction to print in the questions below can be
accomplished either through the `print()` function or by displaying a
value directly.

**TOTAL POINTS: 7**

# Questions

## Q1

(1 POINT)

Print the number of rows in the dataset.

```{r}
# CODE HERE
nrow(dataset)
```

## Q2

(1 POINT)

The rightmost column is the number of rings. Print the maximum number of
rings

```{r}
# CODE HERE
max(dataset$X15)
```

## Q3

(1 POINT)

The leftmost column is the gender with these values: `M`: male, `F`:
female, `I`: infant.

Apply the `filter()` function from tidyverse to select only rows where
gender is infant, and print the number of records.

```{r}
# CODE HERE
records <- dataset %>%
  filter(M =="I")
nrow(records)
```

## Q4

(1 POINT)

Apply the `filter()` function from `tidyverse` to select only rows where
gender is infant or male, and print the number of records.

```{r}
# CODE HERE
records2 <- dataset %>%
  filter(M == "I" | M == "M")
nrow(records2)
```

## Q5

(1 POINT)

Call the `table()` function on the abalone genders to find out how many
of each gender are present.

Print the result.

```{r}
# CODE HERE
table(dataset$M)
```

## Q6

(1 POINT)

Compute the mean value of column 2 (V2) grouped by gender.

V2 is the longest shell measurement.

Requirements: use the `%>%` operator to chain commands, and the
`group_by()` and `summarize()` functions.

```{r}
# CODE HERE
dataset %>%
  group_by(M) %>%
  summarize(mean_val=mean(X0.455))
```

# Q7

(1 POINT)

Compute the MEDIAN value of longest shell measurement for only the
males.

Requirements: use the `%>%` operator to chain commands.

```{r}
# CODE HERE
dataset %>%
  filter(M == "M") %>%
  summarize(median=median(X0.455))
```

# Submission

Save as PDF and upload to Gradescope.
